---
title: "Final Project..."
author: "Ben Essex, Derek Che, & Chris Marston"
date: 04/02/2022
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: pygments
    downcute_theme: "chaos"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

#TODO (add outline comments)

**PROPOSED QUESTIONS** #TODO (we might want to slim these down a bit)

These are the two primary questions we originally set out to answer.

- "Do Covid case rates affect Zoom lecture attendance?"
- "Does weather predict Zoom lecture attendance?"

These are the original questions we had proposed broken down into a more "testable" format as per Scott's instructions.

- "Does weather predict Zoom lecture attendance?"
    - "Does the temperature 2 hours before lecture affect Zoom lecture attendance?"
    - "Does the temperature 1 hour before lecture affect Zoom lecture attendance?"
    - "Does the temperature at the start of lecture affect Zoom lecture attendance?"
    - "Does the type of weather (cloudy, sunny, etc.) affect Zoom lecture attendance?"
- "Do University of Utah Covid case rates affect Zoom lecture attendance?"
    - "Is Zoom lecture attendance affected by the total number of new Covid Cases at the University of Utah?"
    - "Is Zoom lecture attendance affected by the 7-day average of new Covid case rates at the University of Utah?"
- "Do Salt Lake County Covid case rates affect Zoom lecture attendance?"
    - "Is Zoom lecture attendance affected by the SLC County Covid case rates per 100,000?"
    - "Is Zoom lecture attendance affected by the 7-day average case rate per 100,000 in SLC County?"
    - "Is Zoom lecture attendance affected by the 7-day average positive test rate in SLC County?"

## Data Collection

#TODO (add outline comments)

Due to the breadth of scope of the two questions being answered in this study, data was gathered from a range of sources.

**Salt Lake County Covid Data**

- Data was gathered from the Utah Covid Dashboard (overview) website [here](https://coronavirus-dashboard.utah.gov/overview.html) 

**University of Utah Covid Data**

- Data was gathered from the University of Utah Covid information page.
- Important Notes:
    - The U stopped posting Covid data to their public site partway through the collection period.
        - Currently attempting to get the data from them directly...

**Weather Data**

Weather data was pulled from:

- DarkSky for first part of data collection period.
- OpenWeather for last part of data collection period.
- Important Notes:
    - We switched data collection sources part way through the collection period.

**Zoom Attendance Data (headcounts)**

- Gathered directly by attending Zoom and recording attendee count by Zoom.
    - Professor excluded from count.
    - Excluded ourselves if we would have attended in person otherwise.
    - TAs were left in the counts - too difficult to remove them accurately.
    - Attendees counts were taken 10 min into the lecture, at the halfway point, and 10 min before the lecture ended.


## Methodology

#TODO (add outline comments)

## Analysis

```{r}
# import libraries
# data <- read csvs

# +-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
# |  Date |  T+10 | T+40 | T+70 | mean(...) | Temp T-2 | Temp T-1 | Temp T-0 | Weather | Class | UU New Cases | UU 7-Day Avg New Cases | SLCo Case Rate | SLCo 7-Day Avg Case Rate| SLCo 7-Day Avg Positive Test Rate | 
# +-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

# regression analyses:
#   - stimulus: nominal_weather, response: mean_attendance
#     - logistic regression(?)
#   - stimulus: temperature ->
#     - linear regression
#     - T-2 -> mean_attendance
#     - T-1 -> mean_attendance
#     - T-0 -> mean_attendance
#   - stimulus: COVID... ->
#     - linear regression
#     - uu_new_cases -> mean_attendance
#     - uu_new_cases_7day_avg -> mean_attendance
#     - slco_case_rate -> mean_attendance
#     - slco_case_rate_7day_avg -> mean_attendance
#     - slco_pos_test_rate_7day_avg -> mean_attendance
#   - more?

# correlation coefficients
#   - confidence interval
#   - point estimate
```

#TODO (add outline comments)

## Limitations

#TODO (add outline comments)

## Conclusions

#TODO (add outline comments)

